On the number of tiles visited by a\\ line segment on a rectangular grid

### 2022-01-09 paper.tex
Unifying use of $a$ or $b$ in Buffon's problem: always use $a·

The questions studied in this paper are related to Buffon's needle problem and Laplace's extension of it. Buffon's original problem considers tossing a needle of length $\len$ onto ruled paper with parallel lines a distance $b$ apart, and asks what is the probability of the needle crossing a line if it lands in a uniformly random position and orientation. For a needle of length $\len = b$, this is $2/\pi$ (and hence repeated trials of this experiment can be used to estimate $\pi$). This is generalized in \cite{Ramaley69} to $2\len /(\pi b)$ for the expected number of crossings of a needle with arbitrary length.

The Buffon-Laplace needle problem \cite[section 1.1]{Mathai99} considers a needle tossed onto a grid of rectangular tiles of width $a$ and height $b$. The number of visited tiles equals one plus the number of crossings almost surely. The probability of the needle staying within a single tile is computed in \cite{Arnow94} for the case where $\len < \min(a,b)$. One of the problems considered in this paper, as mentioned above, is the complementary question \ref{prob: rand prob} of the probability that the needle visits the maximum number of tiles possible for its length.

@@

\begin{proof}
Suppose first that the grid is simplified to only horizontal lines with spacing $b$. This matches the set-up of Buffon's original needle problem \cite[section 1.1]{Mathai99}, except that here the length $\len$ of the needle may exceed the spacing $b$, allowing it to cross multiple lines. As shown in \cite{Ramaley69}, the expected number of lines crossed equals $2\len/(\pi b)$.
% When $\len<b$ this expectation matches the probability of the needle crossing a grid line, since it cannot cross more than one.

@@

% ***Alex: please check this, as I've done some content changes
Lastly, $\Pr[i \geq 2]$ as computed in Proposition~\ref{prop: Pr i} can be interpreted as the probability of crossing at least one line in Buffon's experiment. Indeed, for $n=2$ the third expression in \eqref{eq: Pr i geq n} reduces again to $2\len/(\pi a)$ for $\len \leq a$, whereas the second reduces to $2\len f(a/\len)/(\pi a)$ for $\len > a$. Using \eqref{eq: f} this is expressed as
\[
\frac 2 \pi \arccos\frac a \len + \frac{2\len}{\pi a} \left(1 - \sqrt{1-\frac{a^2}{\len^2}} \right),
\]
in agreement with the known result for Buffon's needle problem with $\len > a$ \cite[theorem 1.1.2]{Mathai99}. On the other hand, $\Pr[i \geq n]$ for $n \geq 3$ can be seen as the probability of crossing at least one line in a modified version of Buffon's setting in which a needle endpoint is only allowed to move in a region of width $a$ located in the middle of a strip of width $(2n-3)a$.

>>>>

The questions studied in this paper are related to Buffon's needle problem and Laplace's extension of it. Buffon's original problem considers a plane with vertical lines a distance $a$ apart. A needle is placed on the plane with uniformly random position\footnote{It suffices to define position using a horizontal coordinate modulo $a$, for which a uniform distribution can be defined.} and orientation, and the probability of the needle crossing a line is studied. For a needle of length $\len = a$, this probability is $2/\pi$ (and hence repeated trials of this experiment can be used to estimate $\pi$). This is generalized in \cite{Ramaley69} to $2\len /(\pi a)$ for the expected number of crossings of a needle with arbitrary length.

The Buffon-Laplace needle problem \cite[section 1.1]{Mathai99} considers a needle randomly placed on a grid of rectangular tiles of width $a$ and height $b$. The number of visited tiles equals one plus the number of crossings almost surely. The probability of the needle staying within a single tile is computed in \cite{Arnow94} for the case where $\len < \min\{a,b\}$. One of the problems considered in this paper, as mentioned above, is the complementary question \ref{prob: rand prob} of the probability that the needle visits the maximum number of tiles possible for its length.

@@

\begin{proof}
Suppose first that the grid is simplified to only vertical lines with spacing $a$. This matches the set-up of Buffon's original needle problem \cite[section 1.1]{Mathai99}, except that here the length $\len$ of the needle may exceed the spacing $a$, allowing it to cross multiple lines. As shown in \cite{Ramaley69}, the expected number of lines crossed equals $2\len/(\pi a)$.
% When $\len<a$ this expectation matches the probability of the needle crossing a grid line, since it cannot cross more than one.

@@

Lastly, $\Pr[i \geq 2]$ as computed in Proposition~\ref{prop: Pr i} can be interpreted as the probability of crossing at least one line in Buffon's experiment. Indeed, for $n=2$ the third expression in \eqref{eq: Pr i geq n} reduces again to $2\len/(\pi a)$ for $\len \leq a$, whereas the second reduces to $2\len f(a/\len)/(\pi a)$ for $\len > a$. Using \eqref{eq: f} this is expressed as
\[
\frac 2 \pi \arccos\frac a \len + \frac{2\len}{\pi a} \left(1 - \sqrt{1-\frac{a^2}{\len^2}} \right),
\]
in agreement with the known result for Buffon's needle problem with $\len > a$ \cite[theorem 1.1.2]{Mathai99}. On the other hand, $\Pr[i \geq n]$ for $n \geq 3$ can be seen as the probability of crossing at least one line in a modified version of Buffon's setting in which a needle endpoint is only allowed to move in a region of width $a$ located in the middle of a strip of width $(2n-3)a$.
